---
title: "report"
format: html
editor: visual
---

## Background

This GitHub repository reproduces the time series from:

SCHAEFER, DOUGLAS. A.; McDOWELL, WILLIAM H.; SCATENA, FREDRICK N.; ASBURY,CLYDE E. 2000. Effects of hurricane disturbance on stream water concentrations and fluxes in eight tropical forest watersheds of the Luquillo Experimental Forest, Puerto Rico. Journal of Tropical Ecology 16:189±207.

## Data

1)  Download the 4 csv files from the data folder: QuebradaCuenca1-Bisley, QuebradaCuenca2-Bisley, QuebradaCuenca3-Bisley, and RioMameyesPuenteRoto.

2)  Download libraries

```{r}
# | eval: false
library(here)
library(janitor)
library(tidyverse)
library(tidyr)
library(skimr)
library(dplyr)
```

3)  Clean and filter the raw data

```{r}
# Load the data into your environment  
BQ1_data <- read_csv(here("data", "QuebradaCuenca1-Bisley.csv"), na = c("NA"))
BQ2_data <- read_csv(here("data", "QuebradaCuenca2-Bisley.csv"), na = c("NA"))
BQ3_data <- read_csv(here("data", "QuebradaCuenca3-Bisley.csv"), na = c("NA"))
PRM_data <- read_csv(here("data", "RioMameyesPuenteRoto.csv"), , na = c("NA"))



# Cleaning and filtering each data set for necessarily columns (which are the Sample_ID, Sample_Date, concentrations of NO3-N, K, Mg, Ca, NH4-N) 

BQ1_data <- BQ1_data %>% 
  select("Sample_ID", "Sample_Date", "NO3-N", "K", "Mg", "Ca", "NH4-N") %>% 
  clean_names() # changes all the column names to follow the same syntax

BQ2_data <- BQ2_data %>% 
  select("Sample_ID", "Sample_Date", "NO3-N", "K", "Mg", "Ca", "NH4-N") %>% 
  clean_names()

BQ3_data <- BQ3_data %>% 
  select("Sample_ID", "Sample_Date", "NO3-N", "K", "Mg", "Ca", "NH4-N") %>% 
  clean_names()

PRM_data <- PRM_data %>% 
  select("Sample_ID", "Sample_Date", "NO3-N", "K", "Mg", "Ca", "NH4-N") %>% 
  clean_names()

```

## Methods

1)  After downloading and cleaning the data, find the moving average (ma) for all chemical concentrations (K, NO3-N, Mg, Ca, NH4-N) at each site (BQ1, BQ2, BQ3, PRM).

```{r}

# Use the moving average function: 
source(here("outputs", "moving_avg_func.R"))
```

2)  Apply the moving average function to all the data sets for each chemical concentration, and create new columns for each calculated moving average in the existing data frames

```{r}

# (1) ma for all chem concentrations in the BQ1 site 

BQ1_data <- BQ1_data %>% 
  mutate(   # creates new columns in the data frame 
  k_ma = sapply(sample_date,   # the function is applied to all dates in the sample dates 
                moving_avg_func,   # using the moving average function 
                dates = sample_date,   # looks across all sample dates to find the rows that fit the time window 
                conc = k,   # averages the potassium concentrations 
                window_size_wk = 9),   # takes the moving avg with a window of 9 weeks
  no3n_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = no3_n,
                   window_size_wk = 9), 
  mg_ma = sapply(sample_date, 
                 moving_avg_func, 
                 dates = sample_date, 
                 conc = mg,
                 window_size_wk = 9), 
  ca_ma = sapply(sample_date, 
                 moving_avg_func, 
                 dates = sample_date, 
                 conc = ca,
                 window_size_wk = 9), 
  nh4n_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = nh4_n,
                   window_size_wk = 9))


# (2) ma for all chem concentrations in the BQ2 site 

BQ2_data <- BQ2_data %>% 
  mutate(
    k_ma = sapply(sample_date, 
                  moving_avg_func, 
                  dates = sample_date, 
                  conc = k,
                  window_size_wk = 9),
    no3n_ma = sapply(sample_date, 
                     moving_avg_func, 
                     dates = sample_date, 
                     conc = no3_n,
                     window_size_wk = 9), 
    mg_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = mg,
                   window_size_wk = 9), 
    ca_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = ca,
                   window_size_wk = 9), 
    nh4n_ma = sapply(sample_date, 
                     moving_avg_func, 
                     dates = sample_date, 
                     conc = nh4_n,
                     window_size_wk = 9))

# (3) ma for all chem concentrations in the BQ3 site

BQ3_data <- BQ3_data %>% 
  mutate(
    k_ma = sapply(sample_date, 
                  moving_avg_func, 
                  dates = sample_date, 
                  conc = k,
                  window_size_wk = 9),
    no3n_ma = sapply(sample_date, 
                     moving_avg_func, 
                     dates = sample_date, 
                     conc = no3_n,
                     window_size_wk = 9), 
    mg_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = mg,
                   window_size_wk = 9), 
    ca_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = ca,
                   window_size_wk = 9), 
    nh4n_ma = sapply(sample_date, 
                     moving_avg_func, 
                     dates = sample_date, 
                     conc = nh4_n,
                     window_size_wk = 9))

# (4) ma for all chem concentrations in the PRM site

PRM_data <- PRM_data %>% 
  mutate(
    k_ma = sapply(sample_date, 
                  moving_avg_func, 
                  dates = sample_date, 
                  conc = k,
                  window_size_wk = 9),
    no3n_ma = sapply(sample_date, 
                     moving_avg_func, 
                     dates = sample_date, 
                     conc = no3_n,
                     window_size_wk = 9), 
    mg_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = mg,
                   window_size_wk = 9), 
    ca_ma = sapply(sample_date, 
                   moving_avg_func, 
                   dates = sample_date, 
                   conc = ca,
                   window_size_wk = 9), 
    nh4n_ma = sapply(sample_date, 
                     moving_avg_func, 
                     dates = sample_date, 
                     conc = nh4_n,
                     window_size_wk = 9))
```

3)  Combining all the data sets

```{r}
full_data <- rbind(BQ1_data, BQ2_data, BQ3_data, PRM_data)

full_long <- full_data %>% 
  pivot_longer(cols = k_ma:nh4n_ma, 
               names_to = "nutrients", 
               values_to = "concentration")
```

4)  Create the plot

```{r}

```

## Results

## 
